Object-based Auditory Scenes Similarity Retrieval and Classification With Wavelet Scattering

### 2016-07-12 paper/structureScene.tex
revise up to motivations

\IEEEPARstart{T}{he} amount of audio data recorded from our sonic environment has grown considerably over the past decades.
In order to measure the effect of human activity and climate change on animal biodiversity, researchers in the field of ecoacoustics \cite{ECOACOUSTICS2014, krause} have recently undertaken the massive deployment of acoustic sensors throughout the world \cite{warren2006urban, NessSST13, stowell13a, stowell13b}.
In addition, recent work demonstrates the interest of acoustic monitoring for the characterization of human pleasantness in urban areas \cite{lafayPartI, guyot2005urban, ricciardi2015sound} as well as the prediction of annoyance due to traffic \cite{gloaguen}.
Because they bear a strong societal impact and raise many scientific challenges, we believe that these case studies are of considerable interest for the signal processing community.

Unfortunately, the field of signal-based ecoacoustics is still in infancy. \ja{We still haven't strictly defined what ``ecouacoustics'' is. Is there a simple, short definition we can use somewhere before this?}
As a consequence, few well-built datasets are readily available for evaluation purposes.
A closely related, and more mature, field of investigation is the classification of acoustic scenes, where the aim is to predict labels by processing audio data. \ja{So what's the difference between ecoacoustics and acoustic scene classification?}
While narrower in its range of applications, it has the advantage of being rooted in cognitive psychology of categorization \cite{dubois2006cognitive, maffiolo_caracterisation_1999, guastavino_ideal_2006} and has received the attention of the data mining community for some time, resulting in the availability of several well-designed datasets.

Despite the differences, ecoacoustics \ja{?} and acoustic scene classification (ASC) task both raise the following question: how can we efficiently and effectively process a large amount of audio data using representations that are compact, generic in their design, and flexible in their use?

The work described in this paper attempts to answer this question by demonstrating the usefulness of two concepts: 1) using the scattering transform to extract discriminative features subject to a small number of parameters and 2) object-based modeling of the acoustic scene based on unsupervised clustering of each acoustic scene, allowing us to focus the scene description on specific areas of interest. \ja{Again, the last part is a little vague to me.}

@@

One avenue of ecoacoustics research, originating in bioacoustics, attempts to identify the species that produced a specific sonic event of short duration in order to link animal vocalization with behavior\cite{au2008principles, stowell2014large} \ja{Is this the only reason?}.
A more recent paradigm processes the audio stream in a holistic way over larger time scales, without assuming that a single species is present throughout the recording. Automated systems belonging to this paradigm aim to infer global properties of bioacoustic scenes, including biodiversity indices \cite{Bardeli2010}, migration patterns \cite{Obrist2010}, as well as regions of particular interest before human inspection \cite{rosenstock2002landbird,diefenbach2007incorporating} \ja{What does this mean?}.

The same division is found in the field of auditory scene analysis, wherein the interest may either be 1) to isolate and identify precise events, \ie acoustic event detection (AED), or 2) to predict global attributes assigned to large time spans of audio, \ie acoustic scene classification (ASC).
In the latter \ja{What about former? Why are we discussing both if we are going to focus on one? Shouldn't we just say that at some point?}, most state-of-the-art systems more or less follow the bag-of-frames (BOF) approach introduced by Acouturier et al. \cite{aucouturier2007bag}, where a scene is modeled by high-level summary statistics computed from local features.
Being a very generic method, this approach has been applied to many audio classification tasks besides ASC.
Yet the proposed implementation using Gaussian mixture models (GMMs) of mel-frequency cepstral coefficients (MFCCs) has been demonstrated comparably to a mere averaging of the features \cite{lagrange:hal-01082501}.
Indeed, the typical morphology of acoustic scenes is of a ``skeleton of events on a bed of textures'' \cite{nelken_ear_2013}, with a few discrete sound events superimposed upon a stationary acoustic background.
Therefore, it is arguably more beneficial to focus on the large-scale temporal evolution of auditory scenes rather on the joint probability distribution of short-term features.

This statement has some support in auditory psychology as well as sound synthesis based on summary statistics \cite{mcdermott2013summary}. Indeed, studies in cognitive psychology investigating urban sound environment quality pointed out that low-level features such as (perceived or measured) global sound level \ja{Other things like spectral envelope too?}, are not sufficient to fully characterize an acoustic scene \cite{guyot2005urban,kang2006urban}. Rather, it seems that cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness judgment \cite{kuwano_memory_2003} rely upon higher-level cognitive attributes such as the identities of sound sources which compose the scene. \ja{This argument is semantically a little tricky, since technically the low-level features already contain all this high-level information, it's just that the simple models we use to describe them fail to capture this high-level structure} It has been showed that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high-level cognitive classes, \textit{e.g.} the presence of birds are very likely to be heard in parks in urban areas and so are strong pleasantness indicators \cite{lafay:hal-01111782}. Consequently, research in sound perception is now focused the contribution of specific sound sources in the assessment of a sound environment \cite{ricciardi2015sound,lavandier2006contribution}.

Although the complete set of events occuring within a given auditory stream may not be discernable, even to human expects, research has shown that a small set of events, so-called markers, suffice to reliably predict many high-level attributes \cite{lafayPartI}.
For example, the presence of a few bird calls in a smooth traffic hub is likely to be tagged by humans as a park scene in urban areas. \ja{This is very similar to the last few sentences of the last paragraph. What are we trying to say differently here?}

>>>>

\IEEEPARstart{T}{he} amount of audio data recorded from our sonic environment has grown considerably over the past decades.
In order to measure the effect of human activity and climate change on animal biodiversity, researchers in the field of ecoacoustics \cite{ECOACOUSTICS2014, krause} have recently undertaken the massive deployment of acoustic sensors throughout the world \cite{warren2006urban, NessSST13, stowell13a, stowell13b}.

In addition, recent work demonstrates the interest of acoustic monitoring for the characterization of human pleasantness in urban areas \cite{lafayPartI, guyot2005urban, ricciardi2015sound} as well as the prediction of annoyance due to traffic \cite{gloaguen}.
Because they bear a strong societal impact and raise many scientific challenges, we believe that these case studies are of considerable interest for the signal processing community.

Formerly put, ecoacoustics is an interdisciplinary science that investigates natural and anthropogenic sounds and their relationship with the environment over a wide range of study scales at the intersection between ecology, acoustics and computer science. Despite identified needs, the field of signal-based ecoacoustics is still in infancy.
As a consequence, few well-built datasets are readily available for evaluation purposes.
A closely related, and more mature, field of investigation is the classification of acoustic scenes, where the aim is to predict labels by processing audio data.

While narrower in its range of applications, it has the advantage of being rooted in cognitive psychology of categorization \cite{dubois2006cognitive, maffiolo_caracterisation_1999, guastavino_ideal_2006} and has received the attention of the data mining community for some time, resulting in the availability of several well-designed datasets.

Despite the differences, ecoacoustics and acoustic scene classification (ASC) task both raise the following question: how can we efficiently and effectively process a large amount of audio data using representations that are compact, generic in their design, and flexible in their use?

The work described in this paper attempts to answer this question by demonstrating the usefulness of two concepts: 1) using the wavelet scattering transform to extract discriminative features subject to a small number of parameters and 2) object-based modeling of the acoustic scene based on unsupervised clustering of each acoustic scene, allowing us to focus the scene description on specific areas of interest. \ja{Again, the last part is a little vague to me.}

@@

%One avenue of ecoacoustics research, originating in bioacoustics, attempts to identify species properties of vocalizations, for example to link animal vocalization with behavior\cite{au2008principles, stowell2014large}.

A recent paradigm processes the audio stream in a holistic way over larger time scales, without assuming that a single species is present throughout the recording. Automated systems belonging to this paradigm aim to infer global properties of bioacoustic scenes, including biodiversity indices \cite{Bardeli2010}, migration patterns \cite{Obrist2010}, as well as to detect time regions of particular interest requesting detailed human inspection \cite{rosenstock2002landbird,diefenbach2007incorporating}.

%The same division is found in the field of auditory scene analysis, wherein the interest may either be 1) to isolate and identify precise events, \ie acoustic event detection (AED), or 2) to predict global attributes assigned to large time spans of audio, \ie acoustic scene classification (ASC).
In the latter, most state-of-the-art systems more or less follow the bag-of-frames (BOF) approach introduced by Acouturier et al. \cite{aucouturier2007bag}, where a scene is modeled by high-level summary statistics computed from local features.
Being a very generic method, this approach has been applied to many audio classification tasks besides ASC.
Yet the proposed implementation using Gaussian mixture models (GMMs) of mel-frequency cepstral coefficients (MFCCs) has been demonstrated comparably to a mere averaging of the features \cite{lagrange:hal-01082501}.
Indeed, the typical morphology of acoustic scenes is of a ``skeleton of events on a bed of textures'' \cite{nelken2013}, with a few discrete sound events superimposed upon a stationary acoustic background.
Therefore, it is arguably more beneficial to focus on the large-scale temporal evolution of auditory scenes rather on the joint probability distribution of short-term features.

This statement has some support in auditory psychology as well as sound synthesis based on summary statistics \cite{mcdermott2013summary}. Indeed, studies in cognitive psychology investigating urban sound environment quality pointed out that (perceived or measured) global sound level, are not sufficient to fully characterize an acoustic scene \cite{guyot2005urban,kang2006urban}. Rather, it seems that cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness judgment \cite{kuwano_memory_2003} rely upon higher-level cognitive attributes such as the identities of sound sources which compose the scene. \ja{This argument is semantically a little tricky, since technically the low-level features already contain all this high-level information, it's just that the simple models we use to describe them fail to capture this high-level structure} It has been showed that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high-level cognitive classes, \textit{e.g.} the presence of birds are very likely to be heard in parks in urban areas and so are strong pleasantness indicators \cite{lafayPartI}. Consequently, research in sound perception is now strongly focused the contribution of specific sound sources in the assessment of a sound environment \cite{ricciardi2015sound,lavandier2006contribution}. Although the complete set of events occuring within a given auditory stream may not be discernable, even to human expects, research has shown that a small set of events, so-called markers, suffice to reliably predict many high-level attributes.
